Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-08-01 GWLDCI.tex
final readthrough of paper 1

In \cite{wilsonloop} Agarwala and Amat showed that every weakly admissible Wilson loop diagrams can be identified with a type of matroid called a positroid, and observed that this mapping is neither one-to-one nor onto. Each such positroid corresponds to a positroid cell in the appropriate positive Grassmannian, $\Gr(k,n)$ \cite{Postnikov}. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the directions in an associahedron.

@@

A Wilson loop diagram can also be represented visually, which provides useful intuition for many of the results that follow. We depict the diagram $(\cP,V)$ as a circle with marked points, called {\em vertices}; these vertices are labeled by $V$ (preserving the cyclic ordering). The arcs between consecutive vertices are called {\em edges}. There are also $k$ wavy lines in the interior of the diagram depicting the propagators, whose endpoints lie on the edges: specifically, a propagator $p =(i,j)$ has one endpoint on the edge joining vertices $i$ and $i+1$ and another endpoint on the edge joining $j$ and $j+1$, where $i+1$ and $j+1$ denote the successor of $i$ and $j$ respectively in the cyclic order on $V$.

@@

\begin{dfn}\label{admisdfn}
A Wilson loop diagram $W = (\cP,V)$ is {\em admissible} if \begin{enumerate}
\item $|V| \geq |\cP| + 4.$
\item There does not exist a (non-empty) set of propagators $Q \subseteq \cP$ such that $|V(Q)| < |Q| + 3$.
\item There does not exist a pair of propagators $(i_p,j_p),(i_q,j_q) \in \cP$ such that $i_p < i_q < j_p <j_q$.
\end{enumerate}
A Wilson loop diagram is called {\em weakly admissible} if the second and third conditions hold.
 \end{dfn}

The first condition restricts the overall density of propagators in the diagram, while the second imposes an upper bound on how densely the propagators can be fitted into a portion of the diagram. The third ensures that ensures that no propagators cross in the interior of the diagram.

There are a few things to note here.  First, what we call weakly admissible is called admissible in \cite{wilsonloop}.  Many of the results below apply when the hypothesis of admissibility is changed to weak admissibility.

Note also that if two propagators have the same supporting edges, or if a propagator is supported on adjacent edges, then the Wilson loop diagram is not admissible.

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5\})$ is an example of an exact subdiagram; in fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

\begin{dfn}\label{equivdfn} 
Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. We say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and supported on the same vertices, i.e. $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, [n]) = (\cP' \setminus P', [n])$.
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

@@

A subset $A \subseteq E$ is called {\em independent} in $M$ if $A \subseteq B$ for some $B \in \cB$, and {\em dependent} otherwise. (In other words, a basis is simply an independent set of maximal size in $M$.) The {\em rank}  $\rk(A)$ of a subset $A \subseteq E$ is the size of the largest independent set contained in $A$. The rank of the matroid itself is defined to be $\rk(E)$.

@@

Matroid theory relates to the study of Wilson loop diagrams as follows. In \cite{wilsonloop}, Agarwala and Amat show that every weakly admissible Wilson loop diagram with $k$ propagators defines a positroid of rank $k$ on a base set of size $n$, called $M(W)$. In other words, by \cite{Postnikov}, each admissible Wilson loop diagram corresponds to a convex subset of $\Gr(k,n)$, thus connecting the geometry of Wilson loop diagrams with the geometry underlying the Amplituhedron. The independent sets of $M(W)$ can be read directly from the diagram:

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, if $C(W)|_V$ has fewer than $|V|$ rows of non-zero entries then it not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, then $V$ is a dependent set. Any set of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

This section explores the correspondence between Wilson loop diagrams and positroids that was established in \cite{wilsonloop}, (Theorem 3.38). In that paper, the authors show that admissible Wilson loop diagrams define matroids that are also positroids. Furthermore, they show that if two Wilson loop diagrams are equivalent, they define the same matroid(Theorems 1.18 and 3.41).

In this section we develop this analysis further, showing the converse of this result from loc. cit. (see Theorem \ref{same matroid iff equiv}). We also enumerate the number of diagrams that map to an particular matroid in Corollary \ref{number of equiv diagrams}. To do this, we establish a connection between Wilson loop diagrams and polygon dissections in subsection \ref{sec: polygon partitions}, and uncover some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}. This includes the fact that exact subdiagrams correspond to uniform matroids (Theorem \ref{exactuniformthm}), and that the associated matroid of an exact subdiagram can be written as as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a diagonal edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a diagonal connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\emph{Polygon dissection} is a common term in combinatorics for a decomposition of a convex polygon into smaller polygons by means of noncrossing diagonals, see \cite{PSpolygon}.  Each $\tau(W)$ is such an object.

@@

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices and the outer face of $\tau(W)$ is a cycle.  Since $W$ does not admit any propagators of the form $p = (i, i+1)$ due to its admissibility, there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Similarly, there does not exist two propagators $p,q$ such that both $p$ and $q$ start at edge $i$ and end at edge $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge.  Therefore $\tau(W)$ is a simple graph. Finally, the embedding of $\tau(W)$ is induced from the embedding of the graph $W$ and since $W$ is admissible, no pairs of propagators cross. Therefore, it is a planar embedding. 
 \end{proof}

@@

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon dissection $\tau(W)$ has a unique decomposition into maximal triangulated pieces and edges in the polygon of $\tau(W)$.  Call this a \emph{maximal decomposition}.
\end{lem}

@@

\begin{figure}[h]
\bas 
\begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
 \draw (0,0) node {$\bullet$};
\draw (-1,.2) node {$\bullet$};	
\draw (-.5,1.1) node {$\bullet$};
\draw (.8,-.3) node {$\bullet$};	
\draw (1.1,-.5) node {$\bullet$};
\draw(1.1,-.5) -- (.8,-.3) --(0,0)--(-1,.2) -- (-.5,1.1);
\foreach \i in {1,2,...,8} {
      \draw (45*\i:2) node {$\bullet$};
    }
\draw (45*1:2) --(0,0) -- (45*5:2);
\draw (45*6:2) -- (.8,-.3);
\draw (45*8:2) -- (1.1,-.5) -- (45*7:2);
\draw (45*2:2) -- (-.5,1.1) -- (45*3:2);
\draw (45*4:2) -- (-1,.2);
	\end{scope}
	\end{tikzpicture} \qquad \longrightarrow \qquad 
\begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
 \draw (.2,.2) node {$\bullet$};
\draw (.2,-.2) node {$\bullet$};
\draw (-.2,-.2) node {$\bullet$};
\draw (-.2,.2) node {$\bullet$};
\draw (-2,.2) node {$\bullet$};	
\draw (-.5,1.1) node {$\bullet$};
\draw (.8,-.3) node {$\bullet$};	
\draw (1.1,-.5) node {$\bullet$};
\draw(1.1,-.5) -- (.8,-.3) --(.2,-.2); 
\draw (-.2, .2) --(-1,.2) -- (-.5,1.1);
\foreach \i in {1,2,...,8} {
      \draw (45*\i:2) node {$\bullet$};
    }
\draw (45*1:2) --(.2,.2);
\draw (-.2, -.2) -- (45*5:2);
\draw (45*6:2) -- (.8,-.3);
\draw (45*8:2) -- (1.1,-.5) -- (45*7:2);
\draw (45*2:2) -- (-.5,1.1) -- (45*3:2);
\draw (45*4:2) -- (-1,.2);
	\end{scope}
	\end{tikzpicture}
\eas

@@

Next, we introduce the idea of retriangulations. 
\begin{dfn} \label{retriangulation}
  If $W = (\cP, [n])$ and $W' = (\cP', [n])$ are weakly admissible Wilson loop diagrams, then $\tau(W)$ and $\tau(W')$  differ by a \emph{retriangulation} if there is a bijection $f$ between the maximal triangulated pieces of $\tau(W)$ and the maximal triangulated pieces of $\tau(W')$ such that for any maximal triangulated piece $t$ of $\tau(W)$ the vertex sets of $t$ and $f(t)$ are the same.
\end{dfn}
More casually this definition says that the maximal triangulated pieces of $\tau(W)$ and $\tau(W')$ are the same (so far as their vertex sets are concerned) but that some or all of the triangulated pieces may be differently triangulated while the region being triangulated by each piece (which is defined by the vertex set) remains unchanged. This definition of retriangulation makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

The set of propagators $P= \{p,q,r,s \}$ defines an exact subdiagram of $W$, with $V(P) = \{1,2,3,4,5,8,9\}$.  The corresponding subgraph $t$ of $\tau(W)$ has vertex set $\{1, 2,3, 4, 8\}$. Note that we include vertex 3 of $\tau(W)$ in $t$ because the two endpoints of edge 3 in $W$ (namely the vertices 3 and 4 in $W$) are both in $V(P)$. However, we don't include vertex 5 of $\tau(W)$ in $t$because only one of the endpoints of edge 5 in $W$ is in $V(P)$. The same argument applies to vertices 7, 9 and 10. Vertex 6 is excluded because none of its endpoints lie in $V(P)$.

@@

The key point to note is that the subgraph $t$ depends on how the propagators of $P$ sit inside $[n]$, not on how they sit in the subdiagram $(P,V(P))$, which would have outer edges between vertices 8 and 9, 9 and 1, 4 and 5, and 5 and 8.  In particular, $t$ is not the subgraph of $\tau(W)$ consisting only of edges corresponding to propagators of $P$ (as it may also include some of the outer edges of $\tau(W)$), nor is it the polygon dissection of $(P,V(P))$ viewed as a Wilson loop diagram in its own right (which would have vertex set equal to $V(P)$).
\end{eg}

@@

Now suppose $|P|>1$.  Let $j=|P|$, $m= j+3=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, $t$ would have $j -c$ internal edges.

@@

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, we are motivated to examine the matroid properties of exact subdiagrams more closely. In this section, we prove two main results:
\begin{enumerate}
\item The matroid assoicated to an exact subdiagram of $W$ can be written as a contraction of the matroid $M(W)$ by the complementary propagator flat (Theorem \ref{exact diagrams contractions}). 
\item The matroid associated to an exact subdiagram is uniform (Theorem \ref{exactuniformthm} ).
\end{enumerate}

@@

For part (2), suppose that $(P, V(P))$ is a maximal exact subdiagram. By Lemma \ref{lem decompose flat} (1) and Corollary \ref{classifyflats} we can decompose $F(P^c)$ as
\begin{equation}\label{eq decompose F}V(P)^c = F(P^c) = C \sqcup S = F(\Prop(C)) \sqcup S,\end{equation}
where $C$ is the largest cyclic flat contained in $F(P^c)$ and $S$ is an  independent set. We show that $\Prop(C) = P^c$, thus forcing $S$ to be empty. Note that $C$ must be non-empty since $F(P^c)$ is non-empty and not independent (by part (1)).

By Lemma \ref{lem facts about WLD matroids}(2), $\rk(C) = |\Prop(C)|$. Since $S$ is independent, Lemma \ref{lem facts about WLD matroids}(1) gives $|S| = \rk(S)$. By similar arguments to Lemma~\ref{lem decompose flat}, the fact that $S$ is independent and that $C$ is the union of all circuits in $F(P^c)$ implies that every element of $S$ is independent of $C$. Combining these give the first two inequalities below
\ba \rk\big(F(P^c)\big) = \rk(C) + \rk(S)  = |\Prop(C)| + |S| \leq |\Prop(F(P^c))|\;,\label{rank of comp flat} \ea % this equation doesn't really need to be displayed, but it was causing a badbox  
while the inequality comes from the bound on ranks of vertex sets.

Furthermore, from Remark \ref{alt F(P) rmk}, we see that $\Prop (F(P^c)) \subseteq P^c$. Therefore equation \eqref{rank of comp flat} gives \bas |\Prop (C)| + |S| \leq |P^c|  \;.\eas For ease of future notation, write $Q = \Prop(C)^c$. 
Rearranging and adding $|P|$ to both sides gives \ba |P| + |S| \leq |P| + |P^c| - |\Prop(C)| = |Q| \;.\label{P+Sbound}\ea Taking complements of and rearranging equation \eqref{eq decompose F} we have
\bas V(P) \sqcup S = C^c = F(\Prop(C))^c  = V(Q)\;. \eas
Combining this with equation \eqref{P+Sbound} gives 
\ba |V(Q)| = | V(P) | + |S| = | P | + 3 + |S| \leq |Q| +3 \;. \label{Q defines exact}\ea
Since $W$ is admissible this cannot be a strict inequality, so we have $|V(Q)| = |Q| +3$ and the subdiagram $(Q, V(Q))$ is exact.

>>>>

In \cite{wilsonloop} Agarwala and Amat showed that every weakly admissible Wilson loop diagrams can be identified with a type of matroid called a positroid, and observed that this mapping is neither one-to-one nor onto. Each such positroid corresponds to a positroid cell in the appropriate positive Grassmannian $\Gr(k,n)$ \cite{Postnikov}. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the directions in an associahedron.

@@

A Wilson loop diagram can also be represented visually, which provides useful intuition for many of the results that follow. We depict the diagram $(\cP,V)$ as a circle with marked points, called {\em vertices}; these vertices are labeled by $V$ (preserving the cyclic ordering). The arcs between consecutive vertices are called {\em edges}. There are also $k$ wavy lines in the interior of the diagram, each connecting a pair of edges; these depict the propagators. Specifically, a propagator $p =(i,j)$ has one endpoint on the edge joining vertices $i$ and $i+1$ and another endpoint on the edge joining $j$ and $j+1$, where $i+1$ and $j+1$ denote the successor of $i$ and $j$ respectively in the cyclic order on $V$.

@@

\begin{dfn}\label{admisdfn}
A Wilson loop diagram $W = (\cP,V)$ is {\em admissible} if it satisfies the following three conditions: \begin{enumerate}
\item $|V| \geq |\cP| + 4.$
\item There does not exist a (non-empty) set of propagators $Q \subseteq \cP$ such that $|V(Q)| < |Q| + 3$.
\item There does not exist a pair of propagators $(i_p,j_p),(i_q,j_q) \in \cP$ such that $i_p < i_q < j_p <j_q$.
\end{enumerate}
A Wilson loop diagram is called {\em weakly admissible} if the second and third conditions hold.
 \end{dfn}

The first condition restricts the overall density of propagators in the diagram, while the second imposes an upper bound on how densely the propagators can be fitted into any portion of the diagram. The third ensures that ensures that no propagators cross in the interior of the diagram.

There are a few things to note here.  First, what we call weakly admissible here is called admissible in \cite{wilsonloop}.  Many of the results below still apply when the hypothesis of admissibility is changed to weak admissibility.

Note also that if two propagators have the same pair of supporting edges, or if a propagator is supported on two adjacent edges, then the Wilson loop diagram is not admissible or even weakly admissible.

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For the propagator $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5\})$ is an example of an exact subdiagram; in fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

\begin{dfn}\label{equivdfn} 
Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. Let $\sim$ denote the transitive closure of the following relation: $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and supported on the same vertices, i.e. $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, [n]) = (\cP' \setminus P', [n])$.
\end{enumerate}
The relation $\sim$ is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

@@

A subset $A \subseteq E$ is called {\em independent} in $M$ if $A \subseteq B$ for some $B \in \cB$, and is called {\em dependent} otherwise. (In other words, a basis is simply an independent set of maximal size in $M$.) The {\em rank}  $\rk(A)$ of a subset $A \subseteq E$ is the size of the largest independent set contained in $A$. The rank of the matroid itself is defined to be $\rk(E)$.

@@

Matroid theory relates to the study of Wilson loop diagrams as follows. In \cite{wilsonloop}, Agarwala and Marin-Amat show that every weakly admissible Wilson loop diagram $W$ with $k$ propagators defines a positroid $M(W)$ of rank $k$ on a base set of size $n$. Combined with Postnikov's work in \cite{Postnikov}, this shows that each admissible Wilson loop diagram corresponds to a convex subset of $\Gr(k,n)$, thus connecting the geometry of Wilson loop diagrams with the geometry underlying the Amplituhedron.

The independent sets of $M(W)$ can be read directly from the diagram:

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$ we can consider the restriction $C(W)|_V$ of $C(W)$, which consists of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, if $C(W)|_V$ has fewer than $|V|$ rows of non-zero entries then it does not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, then $V$ is a dependent set. Any set of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

This section explores the correspondence between Wilson loop diagrams and positroids that was established in \cite[Theorem 3.38]{wilsonloop}. In that paper, the authors show that admissible Wilson loop diagrams define matroids that are also positroids. Furthermore, they show that if two Wilson loop diagrams are equivalent, then they define the same matroid \cite[Theorems 1.18 and 3.41]{wilsonloop}.

In this section we develop this analysis further, showing the converse of this result from loc. cit. (see Theorem \ref{same matroid iff equiv}). We also enumerate the number of diagrams that map to a particular matroid in Corollary \ref{number of equiv diagrams}. To do this, we establish a connection between Wilson loop diagrams and polygon dissections in subsection \ref{sec: polygon partitions}, and uncover some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}. This includes the fact that exact subdiagrams correspond to uniform matroids (Theorem \ref{exactuniformthm}), and that the associated matroid of an exact subdiagram can be written as as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$ gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a diagonal edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a diagonal connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\emph{Polygon dissection} is a common term in combinatorics for a decomposition of a convex polygon into smaller polygons by means of noncrossing diagonals; see for example \cite{PSpolygon}.  Each $\tau(W)$ is such an object.

@@

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices and the outer face of $\tau(W)$ is a cycle.  Since $W$ does not admit any propagators of the form $p = (i, i+1)$ due to its admissibility, there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Similarly, there does not exist two propagators $p,q$ such that both $p$ and $q$ start at edge $i$ and end at edge $j$. Therefore, no two vertices of $\tau(W)$ can be connected by more than one edge, and so $\tau(W)$ is a simple graph. Finally, the embedding of $\tau(W)$ is induced from the embedding of the graph $W$, and since $W$ is admissible we know that no pair of propagators can cross. Therefore, it is a planar embedding. 
 \end{proof}

@@

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon dissection $\tau(W)$ has a unique decomposition into maximal triangulated pieces and edges in the polygon of $\tau(W)$.  We call this a \emph{maximal decomposition}.
\end{lem}

@@

\begin{figure}[h]
\bas 
\begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
 \draw[gray] (0,0) node {$\bullet$};
\draw[gray] (-1,.2) node {$\bullet$};	
\draw[gray] (-.5,1.1) node {$\bullet$};
\draw[gray] (.8,-.3) node {$\bullet$};	
\draw[gray] (1.1,-.5) node {$\bullet$};
\draw[gray](1.1,-.5) -- (.8,-.3) --(0,0)--(-1,.2) -- (-.5,1.1);
\foreach \i in {1,2,...,8} {
      \draw[gray] (45*\i:2) node {$\bullet$};
    }
\draw[gray] (45*1:2) --(0,0) -- (45*5:2);
\draw[gray] (45*6:2) -- (.8,-.3);
\draw[gray] (45*8:2) -- (1.1,-.5) -- (45*7:2);
\draw[gray] (45*2:2) -- (-.5,1.1) -- (45*3:2);
\draw[gray] (45*4:2) -- (-1,.2);
	\end{scope}
	\end{tikzpicture} \qquad \longrightarrow \qquad 
\begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
 \draw[gray] (.2,.2) node {$\bullet$};
\draw[gray] (.2,-.2) node {$\bullet$};
\draw[gray] (-.2,-.2) node {$\bullet$};
\draw[gray] (-.2,.2) node {$\bullet$};
\draw[gray] (-1,.2) node {$\bullet$};	
\draw[gray] (-.5,1.1) node {$\bullet$};
\draw[gray] (.8,-.3) node {$\bullet$};	
\draw[gray] (1.1,-.5) node {$\bullet$};
\draw[gray](1.1,-.5) -- (.8,-.3) --(.2,-.2); 
\draw[gray] (-.2, .2) --(-1,.2) -- (-.5,1.1);
\foreach \i in {1,2,...,8} {
      \draw[gray] (45*\i:2) node {$\bullet$};
    }
\draw[gray] (45*1:2) --(.2,.2);
\draw[gray] (-.2, -.2) -- (45*5:2);
\draw[gray] (45*6:2) -- (.8,-.3);
\draw[gray] (45*8:2) -- (1.1,-.5) -- (45*7:2);
\draw[gray] (45*2:2) -- (-.5,1.1) -- (45*3:2);
\draw[gray] (45*4:2) -- (-1,.2);
	\end{scope}
	\end{tikzpicture}
\eas

@@

Next, we introduce the idea of retriangulations. 
\begin{dfn} \label{retriangulation}
  If $W = (\cP, [n])$ and $W' = (\cP', [n])$ are weakly admissible Wilson loop diagrams, then $\tau(W)$ and $\tau(W')$  differ by a \emph{retriangulation} if there is a bijection $f$ between the maximal triangulated pieces of $\tau(W)$ and the maximal triangulated pieces of $\tau(W')$ such that for any maximal triangulated piece $t$ of $\tau(W)$ the vertex sets of $t$ and $f(t)$ are the same.
\end{dfn}
More casually this definition says that the maximal triangulated pieces of $\tau(W)$ and $\tau(W')$ are the same so far as their vertex sets are concerned (and hence the regions triangulated by each piece are the same) but that the configuration of triangles within some or all of the triangulated pieces may be different. This definition of retriangulation makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

The set of propagators $P= \{p,q,r,s \}$ defines an exact subdiagram of $W$, with $V(P) = \{1,2,3,4,5,8,9\}$.  The corresponding subgraph $t$ of $\tau(W)$ has vertex set $\{1, 2,3, 4, 8\}$. Note that we include vertex 3 of $\tau(W)$ in $t$ because the two endpoints of edge 3 in $W$ (namely the vertices 3 and 4 in $W$) are both in $V(P)$. However, we don't include vertex 5 of $\tau(W)$ in $t$ because only one of the endpoints of edge 5 in $W$ is in $V(P)$. The same argument applies to vertices 7, 9 and 10. Vertex 6 is excluded because none of its endpoints lie in $V(P)$.

@@

The key point to note is that the subgraph $t$ depends on how the propagators of $P$ sit inside $[n]$, not on how they sit in the subdiagram $(P,V(P))$ (whose corresponding graph would have additional outer edges between vertices 8 and 9, 9 and 1, 4 and 5, and 5 and 8).  In particular, $t$ is not the subgraph of $\tau(W)$ consisting only of edges corresponding to propagators of $P$ (as it may also include some of the outer edges of $\tau(W)$), nor is it the polygon dissection of $(P,V(P))$ viewed as a Wilson loop diagram in its own right (which would have vertex set equal to $V(P)$).
\end{eg}

@@

Now suppose $|P|>1$.  Let $j=|P|$, $m= j+3=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, then $t$ would have $j -c$ internal edges, so we calculate the number of internal edges of $t$.

@@

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, we are motivated to examine the matroid properties of exact subdiagrams more closely. In this section, we prove two main results:
\begin{enumerate}
\item The matroid assoicated to an exact subdiagram of $W$ can be written as a contraction of the matroid $M(W)$ by the complementary propagator flat (Theorem \ref{exact diagrams contractions}). 
\item The matroid associated to an exact subdiagram is uniform (Theorem \ref{exactuniformthm}).
\end{enumerate}

@@

For part (2), suppose that $(P, V(P))$ is a maximal exact subdiagram of $W$. By Lemma \ref{lem decompose flat}(1) and Corollary \ref{classifyflats} we can decompose $F(P^c)$ as
\begin{equation}\label{eq decompose F}V(P)^c = F(P^c) = C \sqcup S = F(\Prop(C)) \sqcup S,\end{equation}
where $C$ is the largest cyclic flat contained in $F(P^c)$ and $S$ is an independent set. We show that $\Prop(C) = P^c$, thus forcing $S$ to be empty and implying that $F(P^c) = C$, i.e. $F(P^c)$ is a cyclic flat. Note that $C$ must be non-empty since $F(P^c)$ is non-empty and not independent (by part (1)).

By similar arguments to Lemma~\ref{lem decompose flat}, the fact that $S$ is independent and that $C$ is the union of all circuits in $F(P^c)$ implies that every element of $S$ is independent of $C$. Further, we know that $\rk(C) = |\Prop(C)|$ from Lemma \ref{lem facts about WLD matroids}(2), and $\rk(S) = |S|$ from Lemma \ref{lem facts about WLD matroids}(1) and the independence of $S$.  Combining these gives
\ba \rk\big(F(P^c)\big) = \rk(C) + \rk(S)  = |\Prop(C)| + |S| \leq |\Prop(F(P^c))|\;,\label{rank of comp flat} \ea 
where the inequality follows from the bound on ranks of vertex sets in Lemma \ref{lem facts about WLD matroids}(1). For ease of future notation, write $Q = \Prop(C)^c$.

From Remark \ref{alt F(P) rmk} it follows that $\Prop (F(P^c)) \subseteq P^c$, and so equation \eqref{rank of comp flat} can be rewritten as \bas |Q^c| + |S| \leq |P^c|  \;.\eas

Rearranging and adding $|P|$ to both sides gives \ba |P| + |S| \leq |P| + |P^c| - |Q^c| = |Q| \;.\label{P+Sbound}\ea 
Meanwhile, by taking complements of equation \eqref{eq decompose F} it follows that
\bas V(P) \sqcup S = C^c = F(\Prop(C))^c  = V(Q)\;. \eas
Combining this with equation \eqref{P+Sbound} gives 
\ba |V(Q)| = | V(P) | + |S| = | P | + 3 + |S| \leq |Q| +3 \;. \label{Q defines exact}\ea
Since $W$ is admissible this cannot be a strict inequality, so we have $|V(Q)| = |Q| +3$ and the subdiagram $(Q, V(Q))$ is exact.
